Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-10-05 2017/ParOptBook/Chapter4.tex
parOptBook: update refs in Chapter4

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
%                           % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%%custom package
%\usepackage{mathtools}
%\usepackage{cite}
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{document}

@@

\emph{Rule 2.} Classify the indices $i, \; 1 \leq i \leq k$, of the trial points from set \eqref{eq4:search_data} according to the number of the problem constraints fulfilled at these points, by constructing the sets
\begin{equation}
  I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
\end{equation}

@@

\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
\begin{gather}
  \label{eq4:characteristic}
  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
  R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),\\
  R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i), \nonumber \\
  \Delta_i=x_i - x_{i-1} \nonumber
\end{gather}
The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm. An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.

@@

\begin{theorem}
  Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions to be satisfied:
  \begin{enumerate}
    \item each domain $Q_j ,1\le j\le m+1$ from \eqref{eq4:q} is a union of a finite number of positive length segments;
    \item each function $g_j ,1\le j\le m+1$ allows Lipschitz extension $G_j (x)$ with the constant $L_j$ over the whole interval $[a,b]$, i.e.,
    \begin{equation}
      g_j(x)=G_j(x),x\in Q_j,1\le j\le m+1;
    \end{equation}
    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ corresponding to the constraints being active at the absolute minimum point $x^*$ (i.e., the constraints, for which $g_\nu(x^*)=0$) satisfy the inequalities
    \begin{equation}
      \label{eq4:reserves}
      0<2\epsilon_\nu<L_\nu(\beta-\alpha)
    \end{equation}
    where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]\subset Q_{m + 1}$ containing the point $x^*$;
    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ corresponding to the constraints which are not active at the absolute minimum point $x^*$ (i.e., $g_\nu(x^* )<0)$ satisfy either inequalities \eqref{eq4:reserves} or the inequalities
    \begin{equation}
      0< \epsilon_\nu <|g_\nu (x^* )|;
    \end{equation}
  \end{enumerate}

  Then:
  \begin{enumerate}
    \item the point $x^*$ is the limit point of the trial sequence $\{x_k \}$ generated by index algorithm for problem \eqref{eq4:problem} at $\epsilon =0$ in the termination condition \eqref{eq4:stop_cond};
    \item any limit point $\overline x$ of the sequence ${x_k }$ is a solution of the problem \eqref{eq4:problem};
    \item the convergence to the limit point $x$ is a bilateral one if $x\not=a$ and $x\not=b$.
    If the conditions \eqref{eq4:reserves} and (4.27) are not satisfied for given vector of reserves $\epsilon_R$ but the
    problem $\eqref{eq4:problem}$ has an $\epsilon$-reserved solution $x_\epsilon$ defined by \eqref{eq4:eps_res_problem}, then:

@@

\textbf{Selection of components of the vector of reserves.} A problem of the choice of concrete component values in the vector of reserves $\epsilon_R$ arises when applying the considered algorithm.
The convergence theorem can help in solving this problem. Condition \eqref{eq4:reserves} can be considered as a recommendation for the selection of the reserve values, for which the property of
convergence to the solution of problem \eqref{ex4:problem} remains. However, in the general case the values of Lipschitz constants are not given \emph{a priori}, and the length of the interval $[\alpha,\beta]$ from $Q_{m + 1}$, including the sought solution $x^*$ is not kwon also. These difficulties are possible to avoid by
substitution of the unknown coefficients $L_j,\: 1\le j\le m$, by their adaptive underestimates $\mu_\nu$ from (4.15) and by assumption that $\beta-\alpha>\delta=\epsilon q$ where $\epsilon >0$ (accuracy) is a parameter defined in the termination condition and the factor $q>1$ can be interpreted as an a priori estimate of how many times the length of the interval $[\alpha,\beta]$ is greater than given accuracy $\epsilon$. As a result, one can take
the reserve estimates in the form (adaptive reserves)
\begin{displaymath}
  \epsilon_\nu=\mu_\nu\delta,\: 1\le\nu\le m.
\end{displaymath}

@@

The problems were solved by the initial index algorithm as well as by the algorithm with $\epsilon$-reservation. The parameters $r_\nu ,\: 1 \le \nu \le 6$, from (4.17) were chosen as
\begin{displaymath}
  r_\nu = \left\{
  \begin{array}{lr}
    \gamma, & k_\nu < 20\\
    2, & k_\nu \ge 20
  \end{array}
  \right.,
\end{displaymath}
where $k_\nu$ is the number of the computed values of the function $g_\nu$ (i.e., the cardinality of the set $I_\nu$). In the test set generated by generator $G_{SH}$ the values $\gamma=20,\: \epsilon = 0.001$ and the reserves $\delta_\nu=0.02,\: 1 \le \nu \le 5$, were used in the experiments. For the problems from the second test set
(generated by $G_{HL}$ ) the values $\gamma = 10, \epsilon=0.0001$ and the parameters $\delta_\nu=0.001,\: 1 \le \nu\le 5$, were used.

@@

\textbf{The constraints’ check order.} he numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
\begin{equation}
  \label{eq4:basic_h}
  H=\{1,2,\dots,m,m+1\},
\end{equation}
will be considered as a base one.
Assume that every particular point $x$ from the search domain $[a,b]$ can have its own order of the problem constraints’ checking. The order of such checking corresponding to the point $x_i$ from (4.12) will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
\begin{equation}
  H(x_i)=\{j_{i1},\dots,j_{im},j_{i,m+1}\},\:0\le i\le k,
\end{equation}
where
\begin{equation}
  j_{i,m+1}=m+1,\:0\le i\le k.
\end{equation}

Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from (4.33), the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$. In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in (4.7) but the index of the last function in the basic numeration (4.32) as the
value of $\nu(x_i )$ for the point $x_i$ , i.e.,
\begin{equation}
  \nu(x_i)=j_{il}.
\end{equation}

\textbf{Algorithm.} The trial selection rules for Fixed Check Order Method (FCOM) already considered above give the base of the new algorithm scheme, which will be called \emph{Adaptive
Check Order Method} (ACOM). Since the alterable constraint check order excludes a possibility to use expression (4.13) to determine the sets $I\nu$, to construct them, an information array connecting each $x_i$ from (4.12) with the indices of all functions, which have been computed in this node is necessary. Hereinafter, the existence of such database is presumed. The rest modifications consist in the following.

@@

\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to (4.35), the index $\nu=\nu(x_i)$ is the basic index from (4.32) corresponding to the value of the last function estimated in this point. These indices are used in all computations based on expressions (4.13)– (4.17), (4.19).

The modifications in the method of using the indices are related to rules (4.17) of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
indices of the boundary points only. Assume
\begin{equation}
  \nu(x_{i-1})=j_{i-1,p}=j_{i,q}\not=\nu(x_i)=j_{i-1,u}=j_{i,x},
\end{equation}
i.e., the problem function with the basic index $\nu(x_{i-1})$ has the index $p$ in the sequence $H(x_{i-1})$ corresponding to the point $x_{i-1}$ and the index $q$ in the sequence $H(x_i)$ corresponding to the point $x_i$. Analogously, the function with the basic index $\nu(x_i)$ corresponds to the index $u$ from the sequence $H(x_{i-1})$ and the index s from the sequence $H(x_i)$. Then, the introduced modification altering the sixth rule of FCOM consists in the following.

If for the indices $p,q, u$ and $s$ from (4.36) the conditions
\begin{equation}
  p<u,q<s,
\end{equation}
are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae (4.17) with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).

Otherwise, when the reverse to (4.37) conditions are fulfilled
\begin{equation}
  p>u,q>s,
\end{equation}
the characteristic $R(i)$ is defined by the third expression from (4.17) with $\nu=\nu(x_{i-1})$ (in this
case, the value $g_\nu(x_i ) is undefined)$. In the cases when conditions (4.37) and (4.38) are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.

@@

\begin{theorem}
Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions are fulfilled:
\begin{enumerate}
  \item each domain $Q_j ,1\le j\le m+1$ from (4.30) represents a union of a finite number of segments with a positive length;
  \item each function $g_j(x): 1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
  \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ satisfy the inequalities
  \begin{equation}
    0<2\epsilon_\nu<L_\nu(\beta-\alpha),
  \end{equation}
  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$, which is the solution of the problem;
  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from (4.15) corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
  \begin{displaymath}
    r_\nu\mu_\nu>2L_\nu,\: 1\le\nu\le m+1,
  \end{displaymath}
  where $r_\nu>1,\: 1\le\nu\le m+1$ are the parameters of the algorithm.
\end{enumerate}

@@

\textbf{The results of the experiments.} Let us consider solving of the problem from Example \ref{ex4:problem} using ACOM with the parameters $r_\nu=2, \epsilon_\nu=0,\: 1\le \nu\le 3$, and $\epsilon=10^{-5}$ in the termination criterion as an illustration. The trial point coordinates executed by the algorithm during the
problem solving practically coincide to the ones marked by three rows of the vertical strokes in Fig. 4.6. However, the number of the computations of the constraint values appears to be different for FCOM and ACOM. In the fixed constraints’ check order the function values have been calculated $k_1 =63,\: k_2 =49,\: k_3 =35$ times while in the adaptive order --- $k_1 =52,\: k_2 =48,\:
k_3 =35$ times. These results are evidence of the decrease of the number of such checks when using ACOM even for the example with two constraints. The efficiency of the proposed scheme reveals itself more clearly when solving the problems with many constraints.

@@

The experiments have been carried out for the algorithms described above with the fixed constraints’ check order and with the adaptive one (FCOM and ACOM, respectively). The parameters $r_\nu, 1\le\nu\le 5$, from (4.17) were defined by the expressions
\begin{displaymath}
  r_\nu = \left\{
  \begin{array}{lr}
    \gamma, & k_\nu < 20\\
    2, & k_\nu \ge 20
  \end{array}
  \right.,
\end{displaymath}
where $k_\nu$ is the number of the calculated values of the function $g_\nu$ (i.e., the cardinality of the set $I_\nu$). In the experiments, the values $\gamma=20,\: \epsilon = 0.001$, and the reserves $\epsilon_\nu=0.1,\: 1\le\nu\le 5$, were set for the set produced by the generator $G_{SH}$. For the problems from the second set (generator $G_{HL}$ ) the values  $\gamma=10,\: \epsilon = 0.0001$, and the reserves $\epsilon_\nu=0.001,\: 1\le\nu\le 5$, were used.

@@

Moreover, the trials executed according to the rules of ACOM appear to be much less computation costly (in average) than the ones executed by FCOM. This fact is illustrated in figures 4.16 and 4.17, where the average number of computations of the function values executed by FCOM and ACOM for given sets are presented. Fig. 4.16 corresponds to the generator $G_{SH}$ while Fig. 4.17 shows the results for the generator $G_{HL}$. In both figures the function indices are plotted on the abscissa axes. The estimates of the average numbers of the computations of these function values during the trials’ execution are marked on the plot axes. The solid lines in the figures connect the points of the estimates corresponding to FCOM while the dashed ones connect the points of estimates corresponding to ACOM. The presented data confirm that introducing the adaptive constraints’ check order can reduce the total number of the computations essentially.

@@

  It follows from this condition, from the rules of the next trial point selection \eqref{eq4:next_point_1_par} and \eqref{eq4:next_point_2_par}, and from the estimate of the value $\mu_\nu$ from \eqref{eq4:lip_est_par}, that
  \begin{equation}
    \label{eq4:4_60}
    \max\{x_t-x^{q+1},x^{q+1}-x_{t-1}\}\le\gamma \Delta_t,
  \end{equation}
  where $\gamma=(r+1)/2r<1$ and
  \begin{equation}
    r=\min\{r_\nu:1\le\nu\le m+1\}>1.
  \end{equation}
  Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.

  At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved. This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not= \varnothing,\: I_{\omega+1} =  \varnothing$. If $I_{m+1}= \varnothing$, the index $\omega=m+1$. If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$. In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
  \begin{displaymath}
    \max(z_{i-1}, z_i)-z^*_\omega\le\mu\Delta_i
  \end{displaymath}
  follows from \eqref{eq4:lip_est_par} and, along with \eqref{eq4:characteristic_par}, gives the estimate
  \begin{displaymath}
    R(i)=\Delta_i\left(1-\frac{\max(z_{i-1},z_i)-z^*_\omega}{r_\omega\mu_\omega\Delta_i}\right)\ge (1-r^{-1})\Delta_i,
  \end{displaymath}
  where $r$ is from (4.61). Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic. Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system (4.54) should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals. Consequently, conditions (4.57) hold.

>>>>

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
%                           % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%%custom package
%\usepackage{mathtools}
%\usepackage{cite}
%\usepackage{amsmath}
%\usepackage{amssymb}
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{document}

@@

\emph{Rule 2.} Classify the indices $i, \; 1 \leq i \leq k$, of the trial points from set \eqref{eq4:search_data} according to the number of the problem constraints fulfilled at these points, by constructing the sets
\begin{equation}
  \label{eq4:13}
  I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
\end{equation}

@@

\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
\begin{gather}
  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
  \label{eq4:characteristic}
  R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),\\
  R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i), \nonumber \\
  \Delta_i=x_i - x_{i-1} \nonumber
\end{gather}
The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm. An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.

@@

\begin{theorem}
  Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions to be satisfied:
  \begin{enumerate}
    \item each domain $Q_j ,1\le j\le m+1$ from \eqref{eq4:q} is a union of a finite number of positive length segments;
    \item each function $g_j ,1\le j\le m+1$ allows Lipschitz extension $G_j (x)$ with the constant $L_j$ over the whole interval $[a,b]$, i.e.,
    \begin{equation}
      g_j(x)=G_j(x),x\in Q_j,1\le j\le m+1;
    \end{equation}
    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ corresponding to the constraints being active at the absolute minimum point $x^*$ (i.e., the constraints, for which $g_\nu(x^*)=0$) satisfy the inequalities
    \begin{equation}
      \label{eq4:reserves}
      0<2\epsilon_\nu<L_\nu(\beta-\alpha)
    \end{equation}
    where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]\subset Q_{m + 1}$ containing the point $x^*$;
    \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ corresponding to the constraints which are not active at the absolute minimum point $x^*$ (i.e., $g_\nu(x^* )<0)$ satisfy either inequalities \eqref{eq4:reserves} or the inequalities
    \begin{equation}
      \label{eq4:27}
      0< \epsilon_\nu <|g_\nu (x^* )|;
    \end{equation}
  \end{enumerate}

  Then:
  \begin{enumerate}
    \item the point $x^*$ is the limit point of the trial sequence $\{x_k \}$ generated by index algorithm for problem \eqref{eq4:problem} at $\epsilon =0$ in the termination condition \eqref{eq4:stop_cond};
    \item any limit point $\overline x$ of the sequence ${x_k }$ is a solution of the problem \eqref{eq4:problem};
    \item the convergence to the limit point $x$ is a bilateral one if $x\not=a$ and $x\not=b$.
    If the conditions \eqref{eq4:reserves} and \eqref{eq4:27} are not satisfied for given vector of reserves $\epsilon_R$ but the
    problem $\eqref{eq4:problem}$ has an $\epsilon$-reserved solution $x_\epsilon$ defined by \eqref{eq4:eps_res_problem}, then:

@@

\textbf{Selection of components of the vector of reserves.} A problem of the choice of concrete component values in the vector of reserves $\epsilon_R$ arises when applying the considered algorithm.
The convergence theorem can help in solving this problem. Condition \eqref{eq4:reserves} can be considered as a recommendation for the selection of the reserve values, for which the property of
convergence to the solution of problem \eqref{ex4:problem} remains. However, in the general case the values of Lipschitz constants are not given \emph{a priori}, and the length of the interval $[\alpha,\beta]$ from $Q_{m + 1}$, including the sought solution $x^*$ is not kwon also. These difficulties are possible to avoid by
substitution of the unknown coefficients $L_j,\: 1\le j\le m$, by their adaptive underestimates $\mu_\nu$ from \eqref{eq4:lip_est} and by assumption that $\beta-\alpha>\delta=\epsilon q$ where $\epsilon >0$ (accuracy) is a parameter defined in the termination condition and the factor $q>1$ can be interpreted as an a priori estimate of how many times the length of the interval $[\alpha,\beta]$ is greater than given accuracy $\epsilon$. As a result, one can take
the reserve estimates in the form (adaptive reserves)
\begin{displaymath}
  \epsilon_\nu=\mu_\nu\delta,\: 1\le\nu\le m.
\end{displaymath}

@@

The problems were solved by the initial index algorithm as well as by the algorithm with $\epsilon$-reservation. The parameters $r_\nu ,\: 1 \le \nu \le 6$, from \eqref{eq4:characteristic} were chosen as
\begin{displaymath}
  r_\nu = \left\{
  \begin{array}{lr}
    \gamma, & k_\nu < 20\\
    2, & k_\nu \ge 20
  \end{array}
  \right.,
\end{displaymath}
where $k_\nu$ is the number of the computed values of the function $g_\nu$ (i.e., the cardinality of the set $I_\nu$). In the test set generated by generator $G_{SH}$ the values $\gamma=20,\: \epsilon = 0.001$ and the reserves $\delta_\nu=0.02,\: 1 \le \nu \le 5$, were used in the experiments. For the problems from the second test set
(generated by $G_{HL}$ ) the values $\gamma = 10, \epsilon=0.0001$ and the parameters $\delta_\nu=0.001,\: 1 \le \nu\le 5$, were used.

@@

\textbf{The constraints’ check order.} he numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
\begin{equation}
  \label{eq4:basic_h}
  H=\{1,2,\dots,m,m+1\},
\end{equation}
will be considered as a base one.
Assume that every particular point $x$ from the search domain $[a,b]$ can have its own order of the problem constraints’ checking. The order of such checking corresponding to the point $x_i$ from \eqref{eq4:characteristic} will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
\begin{equation}
  H(x_i)=\{j_{i1},\dots,j_{im},j_{i,m+1}\},\:0\le i\le k,
\end{equation}
where
\begin{equation}
  j_{i,m+1}=m+1,\:0\le i\le k.
\end{equation}

Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from (4.33), the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$. In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in \eqref{eq4:trial} but the index of the last function in the basic numeration (4.32) as the
value of $\nu(x_i )$ for the point $x_i$ , i.e.,
\begin{equation}
  \nu(x_i)=j_{il}.
\end{equation}

\textbf{Algorithm.} The trial selection rules for Fixed Check Order Method (FCOM) already considered above give the base of the new algorithm scheme, which will be called \emph{Adaptive
Check Order Method} (ACOM). Since the alterable constraint check order excludes a possibility to use expression \eqref{fig:4_13} to determine the sets $I\nu$, to construct them, an information array connecting each $x_i$ from \eqref{eq4:search_data} with the indices of all functions, which have been computed in this node is necessary. Hereinafter, the existence of such database is presumed. The rest modifications consist in the following.

@@

\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to (4.35), the index $\nu=\nu(x_i)$ is the basic index from (4.32) corresponding to the value of the last function estimated in this point. These indices are used in all computations based on expressions \eqref{fig:4_13}-\eqref{eq4:characteristic}, \eqref{eq4:next_point}.

The modifications in the method of using the indices are related to rules \eqref{eq4:characteristic} of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
indices of the boundary points only. Assume
\begin{equation}
  \nu(x_{i-1})=j_{i-1,p}=j_{i,q}\not=\nu(x_i)=j_{i-1,u}=j_{i,x},
\end{equation}
i.e., the problem function with the basic index $\nu(x_{i-1})$ has the index $p$ in the sequence $H(x_{i-1})$ corresponding to the point $x_{i-1}$ and the index $q$ in the sequence $H(x_i)$ corresponding to the point $x_i$. Analogously, the function with the basic index $\nu(x_i)$ corresponds to the index $u$ from the sequence $H(x_{i-1})$ and the index s from the sequence $H(x_i)$. Then, the introduced modification altering the sixth rule of FCOM consists in the following.

If for the indices $p,q, u$ and $s$ from (4.36) the conditions
\begin{equation}
  p<u,q<s,
\end{equation}
are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae \eqref{eq4:characteristic} with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).

Otherwise, when the reverse to (4.37) conditions are fulfilled
\begin{equation}
  p>u,q>s,
\end{equation}
the characteristic $R(i)$ is defined by the third expression from \eqref{eq4:characteristic} with $\nu=\nu(x_{i-1})$ (in this
case, the value $g_\nu(x_i ) is undefined)$. In the cases when conditions (4.37) and (4.38) are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.

@@

\begin{theorem}
Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions are fulfilled:
\begin{enumerate}
  \item each domain $Q_j ,1\le j\le m+1$ from (4.30) represents a union of a finite number of segments with a positive length;
  \item each function $g_j(x): 1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
  \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ satisfy the inequalities
  \begin{equation}
    0<2\epsilon_\nu<L_\nu(\beta-\alpha),
  \end{equation}
  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$, which is the solution of the problem;
  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from \eqref{eq4:lip_est} corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
  \begin{displaymath}
    r_\nu\mu_\nu>2L_\nu,\: 1\le\nu\le m+1,
  \end{displaymath}
  where $r_\nu>1,\: 1\le\nu\le m+1$ are the parameters of the algorithm.
\end{enumerate}

@@

\textbf{The results of the experiments.} Let us consider solving of the problem from Example \ref{ex4:problem} using ACOM with the parameters $r_\nu=2, \epsilon_\nu=0,\: 1\le \nu\le 3$, and $\epsilon=10^{-5}$ in the termination criterion as an illustration. The trial point coordinates executed by the algorithm during the
problem solving practically coincide to the ones marked by three rows of the vertical strokes in Fig. \ref{fig:4_6}. However, the number of the computations of the constraint values appears to be different for FCOM and ACOM. In the fixed constraints’ check order the function values have been calculated $k_1 =63,\: k_2 =49,\: k_3 =35$ times while in the adaptive order --- $k_1 =52,\: k_2 =48,\:
k_3 =35$ times. These results are evidence of the decrease of the number of such checks when using ACOM even for the example with two constraints. The efficiency of the proposed scheme reveals itself more clearly when solving the problems with many constraints.

@@

The experiments have been carried out for the algorithms described above with the fixed constraints’ check order and with the adaptive one (FCOM and ACOM, respectively). The parameters $r_\nu, 1\le\nu\le 5$, from \eqref{eq4:characteristic} were defined by the expressions
\begin{displaymath}
  r_\nu = \left\{
  \begin{array}{lr}
    \gamma, & k_\nu < 20\\
    2, & k_\nu \ge 20
  \end{array}
  \right.,
\end{displaymath}
where $k_\nu$ is the number of the calculated values of the function $g_\nu$ (i.e., the cardinality of the set $I_\nu$). In the experiments, the values $\gamma=20,\: \epsilon = 0.001$, and the reserves $\epsilon_\nu=0.1,\: 1\le\nu\le 5$, were set for the set produced by the generator $G_{SH}$. For the problems from the second set (generator $G_{HL}$ ) the values  $\gamma=10,\: \epsilon = 0.0001$, and the reserves $\epsilon_\nu=0.001,\: 1\le\nu\le 5$, were used.

@@

Moreover, the trials executed according to the rules of ACOM appear to be much less computation costly (in average) than the ones executed by FCOM. This fact is illustrated in figures \ref{fig:4_16} and \ref{fig:4_17}, where the average number of computations of the function values executed by FCOM and ACOM for given sets are presented. Fig. \ref{fig:4_16} corresponds to the generator $G_{SH}$ while Fig. \ref{fig:4_17} shows the results for the generator $G_{HL}$. In both figures the function indices are plotted on the abscissa axes. The estimates of the average numbers of the computations of these function values during the trials’ execution are marked on the plot axes. The solid lines in the figures connect the points of the estimates corresponding to FCOM while the dashed ones connect the points of estimates corresponding to ACOM. The presented data confirm that introducing the adaptive constraints’ check order can reduce the total number of the computations essentially.

@@

  It follows from this condition, from the rules of the next trial point selection \eqref{eq4:next_point_1_par} and \eqref{eq4:next_point_2_par}, and from the estimate of the value $\mu_\nu$ from \eqref{eq4:lip_est_par}, that
  \begin{equation}
    \label{eq4:4_60}
    \max\{x_t-x^{q+1},x^{q+1}-x_{t-1}\}\le\gamma \Delta_t,
  \end{equation}
  where $\gamma=(r+1)/2r<1$ and
  \begin{equation}
    r=\min\{r_\nu:1\le\nu\le m+1\}>1.
  \end{equation}
  \label{eq4:4_61}
  Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.

  At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved. This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not= \varnothing,\: I_{\omega+1} =  \varnothing$. If $I_{m+1}= \varnothing$, the index $\omega=m+1$. If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$. In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
  \begin{displaymath}
    \max(z_{i-1}, z_i)-z^*_\omega\le\mu\Delta_i
  \end{displaymath}
  follows from \eqref{eq4:lip_est_par} and, along with \eqref{eq4:characteristic_par}, gives the estimate
  \begin{displaymath}
    R(i)=\Delta_i\left(1-\frac{\max(z_{i-1},z_i)-z^*_\omega}{r_\omega\mu_\omega\Delta_i}\right)\ge (1-r^{-1})\Delta_i,
  \end{displaymath}
  where $r$ is from \eqref{eq4:4_61}. Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic. Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system (4.54) should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals. Consequently, conditions (4.57) hold.
